---
title: "Statistical Inference Course Project Part 1"
author: "Vineet"
date: "`r format(Sys.Date(), '%d %B, %Y')`"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Part1: Exponential Distribution v/s Central Limit Theorem

### Overview

In this part, we will investigate the exponential distribution in R and compare it with the Central Limit Theorem. The exponential distribution can be simulated in R with rexp(n, lambda) where lambda is the rate parameter. The mean of exponential distribution is 1/lambda and the standard deviation is also 1/lambda. Set lambda = 0.2 for all of the simulations. We will investigate the distribution of averages of 40 exponentials by doing a thousand simulations.

### Simulation

```{r}
library(ggplot2)
sim <- 1000     # no of simulations
n <- 40         # sample size
lambda <- 0.2

df_sim <- as.data.frame(matrix(rexp(n*sim, rate = lambda), sim, n))
df_sim$mean <- rowMeans(df_sim)

```

### Mean Comparison

```{r}

sample_mean <- mean(df_sim$mean)
paste("Sample Mean:", round(sample_mean,4))
theory_mean <- 1/lambda
paste("Theoretical Mean:", theory_mean)

```

The sample and the theoretical means are very close and roughly the same

### Exponential Distribution of Sample Means

```{r}

plt1 <- ggplot(df_sim, aes(x = mean)) +
    geom_histogram(col = "black", alpha = 0.5) +
    geom_vline(aes(xintercept = sample_mean, col = "Sample Mean"), lwd = .8) + 
    geom_vline(aes(xintercept = theory_mean, col = "Theoretical Mean"), lwd = .8) +
    labs(title = "Exponential Distribution of Sample Means", x = "Sample Mean")
plt1

```

We can see above how close both the sample and theoretical means are.

### Variance Comparison

```{r}

sample_var <- var(df_sim$mean)
paste("Sample Variance:", round(sample_var,4))
theory_var <- (1/lambda)^2 / n
paste("Theoretical Variance:", theory_var)

```

The sample and the theoretical variance values are close too

### Showing Distribution of Sample Means as Normal

Based on CLT, the Central Limit Theorem, the averages of samples must follow normal distribution

```{r}

plt2 <- ggplot(df_sim, aes(x = mean)) + 
    geom_histogram(aes(y = ..density..), color = "black", alpha = 0.5) +
    geom_density(col = "green", size = 1) +
    ggtitle("Distribution of Sample means") + xlab("Sample Mean")
plt2

```

The figure shows that the distribution of Sample Means is approximately normal